A Hitchhiker's Guide to Reproducible Research in R

### 2021-07-05 manuscript.Rmd
closes #120
fix typo

One solution to the problem of external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
One comprehensive approach to software management is containerization.
"By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce [@silverSoftwareSimplified2017].
Instead of leaving it to the user which software should be installed how on their computer, **Docker** provides all software used in an analysis, including the operating system itself.
The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
Such a snapshot of the software stack is called an "image".
Packaging all needed software in such an image requires considerable amounts of storage space.
Two major strategies are helping to keep the storage requirements reasonable.
First, there is a community that maintains images for particular purposes.
For example, there are ready-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disc space."))`
Our image includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`.
This recipe is a textual description of all commands that need to be executed to recreate the image.
Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an unambiguous description of how to recreate a given container image.
However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).

@@

To ensure a comprehensive plan, we suggest that researchers borrow from widely employed standards for writing an empirical manuscript.
The introduction and theoretical background provide the basis for the studies design and can be written before gathering data.
We argue that researchers can plan more effectively if they formulate the planned analysis as a method section that again follows best practices for describing empirical research.
To ensure that the analysis is technically feasible, they can translate their method section into executable computer code for the data analysis using simulated data.
The researchers can include a "faked " results section based on the simulated data, employing dynamic document generation.
Repeated simulation with varying parameters, a so-called Monte Carlo simulation, can then be employed to plan resources to assure efficiency.
For example, researchers can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study.

>>>>

One solution to the problem of external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
One comprehensive approach to software management is containerization.
"By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce [@silverSoftwareSimplified2017].
Instead of leaving it to the user which software should be installed how on their computer, **Docker** provides all software used in an analysis, including the operating system itself.
The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
Such a snapshot of the software stack is called an "image".
Packaging all needed software in such an image requires considerable amounts of storage space.
Two major strategies are helping to keep the storage requirements reasonable.
First, there is a community that maintains images for particular purposes.
For example, there are ready-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disc space."))`
Our image includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`.
This recipe is a textual description of all commands that need to be executed to recreate the image.
Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an unambiguous description of how to recreate a given container image.
However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
A more comprehensive overview over how to employ containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.

@@

To ensure a comprehensive plan, we suggest that researchers borrow from widely employed standards for writing an empirical manuscript.
The introduction and theoretical background provide the basis for the studies design and can be written before gathering data.
We argue that researchers can plan more effectively if they formulate the planned analysis as a method section that again follows best practices for describing empirical research.
To ensure that the analysis is technically feasible, they can translate their method section into executable computer code for the data analysis using simulated data.
The researchers can include a "faked" results section based on the simulated data, employing dynamic document generation.
Repeated simulation with varying parameters, a so-called Monte Carlo simulation, can then be employed to plan resources to assure efficiency.
For example, researchers can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study.
